SSA based global code motion of identical computations

### 2017-04-14 hoist.tex
remove pessimistic

\GCC{} recently got code-hoisting \cite{GCCCodeHoisting} which is implemented as
part of \GVN{}-\PRE{}: it uses the set of ANTIC\_IN and AVAIL\_OUT value
expressions computed for \PRE{}. ANTIC\_IN[B] contains very busy expressions at
basic block B i.e., values computed on all paths from B to exit and
AVAIL\_OUT[B] contains values which are already available. The algorithm hoists
top down to a predecessor.  It uses to know what expressions will be computed on
every path from B to exit, and can be computed in B: that is the safety
condition.  It uses AVAIL\_OUT[B] to subtract out those values already being
computed.  The cost function is: for each hoist candidate, if all successors of
B are dominated by B, then we know insertion into B will eliminate all the
remaining computations.  It then checks to see if at least one successor of B
has the value available.  This avoids hoisting it way up the chain to ANTIC.  It
also checks to ensure that B has multiple successors, since hoisting in a
straight line is pointless.  The algorithm continues top down the dominator
tree, iterating with \PRE{} until no more changes.  One advantage of \GCC{}
implementation is that it works in sync with the \GVN{}-\PRE{} such that when
new hoisting opportunities are created by \GVN{}-\PRE{}, code-hoisting will
hoist them.

@@

The process of computing \GVN{} can be on-demand (as we come across an instruction)
or, precomputed (computing \GVN{} of all the instructions beforehand). Which
process to choose is determined by the scope of code-hoisting we want to
perform. In a pessimistic approach, as described in
Section~\ref{subsec:pessimistic}, we want to hoist a limited set of instructions
from the sibling branches as we iterate the DFS tree bottom-up, it is sufficient
to compute \GVN{} values on-demand. Whereas, in the optimistic approach, as
described in Section~\ref{subsec:optimistic}, we want to hoist as many
instructions as possible, and it would require \GVN{} values to be precomputed.

@@

\subsection{Legality check}
\label{subsec:legality}
Since the equality of candidates is purely based on the value they compute, we
need to establish if hoisting them to a common dominator would be feasible. Once
a common dominator is found, we check whether all the use-operands of the set of
instructions are available at that position. In some cases when the operands are not
available, it is possible to reinstantiate (remateralize) the use-operands, thus
passing the legality check.

Subsequently, it is checked that the side-effects of the computations does not
intersect with any side-effects between the instructions to be hoisted and their
hoisting point. For memory operations like loads, stores, calls etc., it is also
required to check that all the paths from the hoisting point to the end of the
function should execute the exact instruction, in order to guarantee
correctness.

Moreover, hoisting of memory operations is tricky on paths which have indirect
branch targets e.g., landing pad, case statements, goto labels etc., because it
becomes difficult to prove that all the paths from hoisting point to the end of
the function would execute the instruction. In our current implementation we
discard hoisting through such paths.

In the optimistic approach, described in Section~\ref{subsec:optimistic}, it is
possible that a common hoisting point of all the instructions is either too far
away, or not legally possible. In these cases, it is still possible to
`partially' hoist a subset of instructions by splitting the set of candidates
and finding a closer hoisting point for each subset. For more details see
Section~\ref{subsec:partition}.

>>>>

\GCC{} recently got code-hoisting \cite{GCCCodeHoisting} which is implemented as
part of \GVN{}-\PRE{}: it uses the set of ANTIC\_IN and AVAIL\_OUT value
expressions computed for \PRE{}. ANTIC\_IN[B] contains very busy expressions at
basic block B i.e., values computed on all paths from B to exit and
AVAIL\_OUT[B] contains values which are already available. The algorithm hoists
top down to a predecessor.  It uses ANTIC\_IN[B] to know what expressions will
be computed on every path from B to exit, and can be computed in B.  It uses
AVAIL\_OUT[B] to subtract out those values already being computed.  The cost
function is: for each hoist candidate, if all successors of B are dominated by
B, then we know insertion into B will eliminate all the remaining computations.
It then checks to see if at least one successor of B has the value available.
This avoids hoisting it way up the chain to ANTIC.  It also checks to ensure
that B has multiple successors, since hoisting in a straight line is pointless.
The algorithm continues top down the dominator tree, iterating with \PRE{} until
no more changes.  One advantage of \GCC{} implementation is that it works in
sync with the \GVN{}-\PRE{} such that when new hoisting opportunities are
created by \GVN{}-\PRE{}, code-hoisting will hoist them.

@@

The process of computing \GVN{} can be on-demand (as we come across an
instruction) or, precomputed (computing \GVN{} of all the instructions
beforehand). Which process to choose is determined by the scope of code-hoisting
we want to perform. In a pessimistic approach, we want to hoist a limited set of
instructions from the sibling branches as we iterate the DFS tree bottom-up, it
is sufficient to compute \GVN{} values on-demand. Whereas, in the optimistic
approach, as described in Section~\ref{subsec:optimistic}, we want to move as
many instructions as possible, and it would require \GVN{} values to be
precomputed.

@@

\subsection{Legality check}
\label{subsec:legality}
Since the equality of candidates is purely based on the value they compute, we
need to establish if hoisting them to a common dominator or sinking them to a
common post-dominator would be feasible. Once a common dominator
(post-dominator) is found, we check whether all the use-operands of the set of
instructions are available at that position. In some cases when the operands are
not available, it is possible to reinstantiate (remateralize) the use-operands,
thus passing the legality check.

Subsequently, it is checked that the side-effects of the computations does not
intersect with any side-effects between the instructions to be hoisted/sunk and
their hoisting/sinking point.

Moreover, moving memory operations is tricky on paths which have indirect
branch targets e.g., landing pad, case statements, goto labels etc., because it
becomes difficult to prove safety checks. In our current implementation we
discard candidates on those paths.

In the approach, described in Section~\ref{subsec:optimistic}, it is possible
that a common hoisting point of all the instructions is either too far away, or
not legally possible. In these cases, it is still possible to `partially' hoist
a subset of instructions by splitting the set of candidates and finding a closer
hoisting point for each subset. For more details see
Section~\ref{subsec:partition}.
